---
title: "GLMM"
author: "Jonathan Zhang"
date: "4/6/2020"
output: html_document
---

###Setup
```{r, echo = T, results = 'hide', message = FALSE}
library(tidyverse)
library(stats)
library(lme4)
dat <- readRDS("dat2.rds")
```
###Generalized Mixed Model
We use a log linear mixed model to model the number of new cases.

$log(E[Y_{ij}|b_i])=\beta_0+\beta_1 Time_{ij}+\beta_2 Time_{ij}^2+\beta_3 GHSScore_i+\beta_4 AgeGEQ65_i+\beta_5 UrbanPop_i+b_{i0}+b_{i1}Time_{ij}$

We could consider another random effect for $Time_{ij}^2$ as well.

We remove NAs that could arise from missing data on certain countries, or from inconsistent naming.
```{r}
dat <- dat %>% mutate(day2 = day^2) %>% drop_na(AgeGEQ65) %>% drop_na(UrbanPop)
```

We fit the previous model, with and without the $Time_{ij}^2$ random effect term with glmer

```{r}
glmm1 <- glmer(new_cases ~ day + day2 + GHS_Score + AgeGEQ65 + UrbanPop + (day | Country.Region), data = dat, family = poisson)
glmm2 <- glmer(new_cases ~ day + day2 + GHS_Score + AgeGEQ65 + UrbanPop + (day + day2 | Country.Region), data = dat, family = poisson)
```

There are a couple of warnings here. The first deals with non-convergence, which could be caused by having multiple random and fixed effects. I tried using both bobyqa and Nelder-Mead for the optimzation algorithm in glmer but could not achieve convergence at their threshold. A model with only two random effects (intercept and day) and two fixed effects (intercept and day) does not have this error. The second warning deals with unidentifiability. I have tried rescaling parameters by normalizing them and using the log odds from percentiles but could not get this warning to go away either. Nonetheless, we still get a fit.

```{r}
dats <- dat
dats["AgeGEQ65"] <- scale(dats["AgeGEQ65"])
dats["GHS_Score"] <- scale(dats["GHS_Score"])
glmm11 <- glmer(new_cases ~ day + GHS_Score + AgeGEQ65 + UrbanPop + (day | Country.Region), data = dats, family = poisson)
```

```{r}
summary(glmm1)
```

For Model 1, it seems that for fixed effects, all but AgeGEQ65 and UrbanPop were significant.

```{r}
summary(glmm2)
```

For Model 2, it seems that for fixed effects, all but GHS_Score, AgeGEQ65 and UrbanPop were significant.

We now compare the model to the observed counts for several different countries and view their predictions.

###US Model 1

```{r}
country = 103
country_name = 'US'
rand_ef = ranef(glmm1)$Country.Region[country,]
fix_ef = fixef(glmm1)
coef = c(fix_ef[[1]]+rand_ef[[1]],fix_ef[[2]]+rand_ef[[2]],fix_ef[[3]],fix_ef[[4]],fix_ef[[5]],fix_ef[[6]])
dat2 <- dat %>% 
  filter(Country.Region==country_name) %>% 
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=dat2)+geom_line(aes(x=day,y=new_cases))+geom_line(aes(x=day,y=model))
```

and the prediction

```{r}
t<-tibble(day=1:100)%>%mutate(day2=(1:100)^2,GHS_Score=dat2$GHS_Score[1],AgeGEQ65=dat2$AgeGEQ65[1],UrbanPop=dat2$UrbanPop[1])%>%
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=t)+geom_line(aes(x=day,y=model))
```

###US Model 2

```{r}
country = 103
country_name = 'US'
rand_ef = ranef(glmm2)$Country.Region[country,]
fix_ef = fixef(glmm2)
coef = c(fix_ef[[1]]+rand_ef[[1]],fix_ef[[2]]+rand_ef[[2]],fix_ef[[3]]+rand_ef[[3]],fix_ef[[4]],fix_ef[[5]],fix_ef[[6]])
dat2 <- dat %>% 
  filter(Country.Region==country_name) %>% 
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=dat2)+geom_line(aes(x=day,y=new_cases))+geom_line(aes(x=day,y=model))
```

and the prediction

```{r}
t<-tibble(day=1:100)%>%mutate(day2=(1:100)^2,GHS_Score=dat2$GHS_Score[1],AgeGEQ65=dat2$AgeGEQ65[1],UrbanPop=dat2$UrbanPop[1])%>%
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=t)+geom_line(aes(x=day,y=model))
```

###China Model 1

```{r}
country = 23
country_name = 'China'
rand_ef = ranef(glmm1)$Country.Region[country,]
fix_ef = fixef(glmm1)
coef = c(fix_ef[[1]]+rand_ef[[1]],fix_ef[[2]]+rand_ef[[2]],fix_ef[[3]],fix_ef[[4]],fix_ef[[5]],fix_ef[[6]])
dat2 <- dat %>% 
  filter(Country.Region==country_name) %>% 
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=dat2)+geom_line(aes(x=day,y=new_cases))+geom_line(aes(x=day,y=model))
```

and the prediction

```{r}
t<-tibble(day=1:100)%>%mutate(day2=(1:100)^2,GHS_Score=dat2$GHS_Score[1],AgeGEQ65=dat2$AgeGEQ65[1],UrbanPop=dat2$UrbanPop[1])%>%
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=t)+geom_line(aes(x=day,y=model))
```

###China Model 2

```{r}
country = 23
country_name = 'China'
rand_ef = ranef(glmm2)$Country.Region[country,]
fix_ef = fixef(glmm2)
coef = c(fix_ef[[1]]+rand_ef[[1]],fix_ef[[2]]+rand_ef[[2]],fix_ef[[3]]+rand_ef[[3]],fix_ef[[4]],fix_ef[[5]],fix_ef[[6]])
dat2 <- dat %>% 
  filter(Country.Region==country_name) %>% 
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=dat2)+geom_line(aes(x=day,y=new_cases))+geom_line(aes(x=day,y=model))
```

and the prediction

```{r}
t<-tibble(day=1:100)%>%mutate(day2=(1:100)^2,GHS_Score=dat2$GHS_Score[1],AgeGEQ65=dat2$AgeGEQ65[1],UrbanPop=dat2$UrbanPop[1])%>%
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=t)+geom_line(aes(x=day,y=model))
```

###South Korea Model 1

```{r}
country = 54
country_name = 'Korea, South'
rand_ef = ranef(glmm1)$Country.Region[country,]
fix_ef = fixef(glmm1)
coef = c(fix_ef[[1]]+rand_ef[[1]],fix_ef[[2]]+rand_ef[[2]],fix_ef[[3]],fix_ef[[4]],fix_ef[[5]],fix_ef[[6]])
dat2 <- dat %>% 
  filter(Country.Region==country_name) %>% 
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=dat2)+geom_line(aes(x=day,y=new_cases))+geom_line(aes(x=day,y=model))
```

and the prediction

```{r}
t<-tibble(day=1:100)%>%mutate(day2=(1:100)^2,GHS_Score=dat2$GHS_Score[1],AgeGEQ65=dat2$AgeGEQ65[1],UrbanPop=dat2$UrbanPop[1])%>%
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=t)+geom_line(aes(x=day,y=model))
```

###South Korea Model 2

```{r}
country = 54
country_name = 'Korea, South'
rand_ef = ranef(glmm2)$Country.Region[country,]
fix_ef = fixef(glmm2)
coef = c(fix_ef[[1]]+rand_ef[[1]],fix_ef[[2]]+rand_ef[[2]],fix_ef[[3]]+rand_ef[[3]],fix_ef[[4]],fix_ef[[5]],fix_ef[[6]])
dat2 <- dat %>% 
  filter(Country.Region==country_name) %>% 
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=dat2)+geom_line(aes(x=day,y=new_cases))+geom_line(aes(x=day,y=model))
```

and the prediction

```{r}
t<-tibble(day=1:100)%>%mutate(day2=(1:100)^2,GHS_Score=dat2$GHS_Score[1],AgeGEQ65=dat2$AgeGEQ65[1],UrbanPop=dat2$UrbanPop[1])%>%
  mutate(model=exp(coef[1]+coef[2]*day+coef[3]*day^2+coef[[4]]*GHS_Score+fix_ef[[5]]*AgeGEQ65+fix_ef[[6]]*UrbanPop))
ggplot(data=t)+geom_line(aes(x=day,y=model))
```

