---
title: "NLME"
author: "Kushal Shah"
date: "4/5/2020"
output: html_document
---

### Setup and Read-in

```{r, echo = T, results = 'hide', message = FALSE}
library(tidyverse)
library(stats)
library(lme4)
library(nlme)

dat <- readRDS("dat.rds")
dat_old <- readRDS("dat_old.rds")
```

### Bell Curve Functional Form

We begin with the assumption that the normal distribution kernel, scaled by a constant, appears to be a good approximation. Here, we have three parameters that can be customized to every country: d (constant that changes the height of the curve), mu (approximately represents the day of peak cases), and var (spread of the bell curve).

```{r}
# Functional form of new COVID cases
bellcurve.model <- function(d, mu, var, x) {
  f <- d*exp(-((x - mu)^2) / (2*var))
  return(f)
}
```

## Modeling the Old Dataset

Using data from approximately ten days ago, called `dat_old`.

### Nonlinear Mixed Effects Model (Old Dataset)

Uses the nlme function of the nlme library to fit a nonlinear mixed model. Here, every country is given a random effect for d, mu, and var. We start at fixed estimates of d=1000, mu=20, var=20.

The model updates following the base model are something I don't fully understand yet, but each update increase the accuracy of the model.

```{r}
# NLME model using old dataset
baseModel <- nlme(new_cases ~ bellcurve.model(d, mu, var, x = day), 
                  data = dat_old, 
                  fixed = list(d ~ 1, mu ~ 1, var ~ 1),
                  random = d + mu + var ~ 1|Country.Region,
                  start = list(fixed = c(1000, 20, 20)),
                  na.action = na.omit)

# Updates to increase accuracy of the model
nestedModel <- update(baseModel, fixed = list(d ~ 1, mu ~ 1, var ~ 1), start = fixef(baseModel))
nestedModel2 <- update(nestedModel, fixed = list(d ~ 1, mu ~ 1, var ~ 1), start = fixef(nestedModel))
```

### Model Summary

#### Fixed Effects

```{r}
# Fixed effects of the model parameters
summary(nestedModel2)
```

#### Random Effects

```{r}
# Random effects of the model parameters by country
ranef(nestedModel2)
```

### Model Fit and Predictions for USA

From the following output, it seems like the model had good predictions on the USA data and accurately has the peak of the bell curve not occuring well past the latest day available in the dataset.

```{r}
# Data manipulation to create USA predictions in the next step
usa <- filter(dat_old, Country.Region == "US")
preds <- predict(nestedModel2)
preds_dat <- data.frame(country = attr(preds, "names"), pred = preds)
preds_dat_usa <- preds_dat %>% 
  filter(country == "US") 
preds_dat_usa <- preds_dat_usa %>%
  mutate(day = as.numeric(rownames(preds_dat_usa)) - 1)

# Model has great predictions on USA data up till March 25, predicting a normal curve with peak new cases at day 29
plot(usa$day, usa$new_cases)
lines(preds_dat_usa$day, preds_dat_usa$pred)
```

## Modeling the New Dataset

### Nonlinear Mixed Effects Model (NLME)

Now we try to run the same model using the new data, called `dat`. If you uncomment this code and run it, you'll get an error: `Error in nlme.formula(new_cases ~ bellcurve.model(d, mu, var, x = day),  : Singularity in backsolve at level 0, block 1`

```{r}
#baseModel2 <- nlme(new_cases ~ bellcurve.model(d, mu, var, x = day), 
#                  data = dat, 
#                  fixed = list(d ~ 1, mu ~ 1, var ~ 1),
#                  random = d + mu + var ~ 1|Country.Region,
#                  start = list(fixed = c(1000, 20, 20)),
#                  na.action = na.omit)
```

### Nonlinear Least Squares

This is an alternative modeling method we can use, but it would just be limited to one country at at time without borrowing information from other countries. Also probably limited to countries with lots of data. Here's a quick example.

```{r}
# Same bell curve function as before
bellcurve.model <- function(d, mu, var, x) {
  f <- d*exp(-((x - mu)^2) / (2*var))
  return(f)
}
```


#### South Korea

Starting parameter estimates chosen as d=800, mu=15, and var=25 based on a quick look at South Korea's scatterplot.

```{r}
southkorea <- filter(dat, Country.Region == "Korea, South")

m1 <- nls(new_cases ~ bellcurve.model(d, mu, var, x = day), 
          start = list(d = 800, mu = 15, var = 25), 
          data = southkorea, 
          #trace = TRUE, 
          lower = list(d = 1, mu = 1, var = 0))

plot(southkorea$day, southkorea$new_cases)
lines(southkorea$day, predict(m1))

# South Korea estimates from the model
summary(m1)
```

#### USA

Starting parameter estimates chosen as d=5000, mu=15, and var=25 based on a looking at USA's scatterplot.

```{r}
usa <- filter(dat, Country.Region == "US")

# Model fit, with lower bounds for parameter estimates that can be specified for the port algorithm
m2 <- nls(new_cases ~ bellcurve.model(d, mu, var, x = day), 
          start = list(d = 5000, mu = 15, var = 25), 
          data = usa, 
          #trace = TRUE, 
          algorithm = "port", 
          lower = list(d = 1, mu = 1, var = 0))

plot(usa$day, usa$new_cases)
lines(usa$day, predict(m2))

# USA estimates from the model
summary(m2)
```

