---
title: "prophet"
author: "Emily Damone"
date: "April 13, 2020"
output: html_document
---

US forecast of total cases using logistic growth - need to add 
```{r setup, include=FALSE}
library(tidyverse)
library(prophet)
library(caret)
library(randomForest)

```

```{r}
 dat <- readRDS("dat3.rds")
# dat_US <- dat[dat$Country.Region == "US", ]
# prophet_dat_US <- as.data.frame(dat_US$date)
# prophet_dat_US$y <- as.numeric(dat_US$total_cases)
# #prophet needs a capacity when forecasting logistic, currently total population - we would need this for all countries
# #prophet_dat$cap <- rep(328200000, length(prophet_dat$y))
# prophet_dat_US$cap <- as.numeric(dat_US$TotalPop)
# 
# 
# colnames(prophet_dat_US) <- c("ds", "y", "cap")
# 
# now_US <- prophet(prophet_dat_US, growth = "logistic")
# 
# future_US <- make_future_dataframe(now_US, periods = 10)
# future_US$cap <- rep(prophet_dat_US[1, 3], length(future$ds))
# 
# forecast_US <- predict(now_US, future_US, )
# plot(now_US, forecast_US, plot_cap=FALSE, uncertainty = TRUE)

```

```{r}
# dat_china <- dat[dat$Country.Region == "China", ]
# prophet_dat_china <- as.data.frame(dat_china$date)
# prophet_dat_china$y <- as.numeric(dat_china$total_cases)
# prophet_dat_china$cap <- as.numeric(dat_china$TotalPop)
# 
# 
# colnames(prophet_dat_china) <- c("ds", "y", "cap")
# 
# now_china <- prophet(prophet_dat_china, growth = "logistic")
# 
# future_china <- make_future_dataframe(now_china, periods = 10)
# future_china$cap <- rep(prophet_dat_china[1, 3], length(future_china$ds))
# 
# forecast_china <- predict(now_china, future_china, )
# plot(now_china, forecast_china, plot_cap=FALSE, uncertainty = TRUE)

forecast <- function(country, data, numPred){
  dat <- data[data$Country.Region == country, ]
  prophet_dat <- as.data.frame(dat$date)
  prophet_dat$y <- as.numeric(dat$total_cases)
  prophet_dat$cap <- as.numeric(dat$TotalPop)
  
  colnames(prophet_dat) <- c("ds", "y", "cap")
  
  now <- prophet(prophet_dat, growth= "logistic")
  future <- make_future_dataframe(now, periods=numPred)
  future$cap <- rep(prophet_dat[1,3], length(future$ds))
  
  forecast <- predict(now, future, )
  print(plot(now, forecast, plot_cap=F, uncertainty = T, ylabel = country) +  add_changepoints_to_plot(now))
  return(list(now,forecast))
}

US_forecast <- forecast("US", dat, 10)
Korea_forecast <- forecast("Korea, South", dat, 10)
Italy <- forecast("Italy", dat, 10)
China <- forecast("China", dat, 10)
```

So I adjusted the cap to be 1% of the population:
```{r}
forecast_1 <- function(country, data, numPred){
  dat <- data[data$Country.Region == country, ]
  prophet_dat <- as.data.frame(dat$date)
  prophet_dat$y <- as.numeric(dat$total_cases)
  prophet_dat$cap <- as.numeric(dat$TotalPop*.01)
  
  colnames(prophet_dat) <- c("ds", "y", "cap")
  
  now <- prophet(prophet_dat, growth= "logistic")
  future <- make_future_dataframe(now, periods=numPred)
  future$cap <- rep(prophet_dat[1,3], length(future$ds))
  
  forecast <- predict(now, future, )
  print(plot(now, forecast, plot_cap=F, uncertainty = T, ylabel = country) +  add_changepoints_to_plot(now))
  return(list(now,forecast))
}

US_forecast_1 <- forecast_1("US", dat, 10)
Korea_forecast_1 <- forecast_1("Korea, South", dat, 10)
Italy_1 <- forecast_1("Italy", dat, 10)
China_1 <- forecast_1("China", dat, 10)
```

The logistic model is pretty bad when there is actual flattening of the curve, it wants to estimate higher (even when just fitting not predicting) - I think this is mainly due to the cap being MUCH higher than where the cases actually levelled off. Could use this and say this would be fine to predict what would happen if we did nothing - currently what US tracks to is very close to the pure logistic curve whereas when flattening the curve doesn't match the model at all.

So here it is with linear: 
```{r}
forecast_linear <- function(country, data, numPred){
  dat <- data[data$Country.Region == country, ]
  prophet_dat <- as.data.frame(dat$date)
  prophet_dat$y <- as.numeric(dat$total_cases)
  
  colnames(prophet_dat) <- c("ds", "y")
  
  now <- prophet(prophet_dat, growth= "linear")
  future <- make_future_dataframe(now, periods=numPred)
  
  forecast <- predict(now, future, )
  print(plot(now, forecast, plot_cap=F, uncertainty = T, ylabel = country) +  add_changepoints_to_plot(now))
  return(list(now,forecast))
}

linear_US <- forecast_linear("US", dat, 10)
linear_Korea <- forecast_linear("Korea, South", dat, 10)
linear_Italy <- forecast_linear("Italy", dat, 10)
linear_China <- forecast_linear("China", dat, 10)
```

Attempting a Random Forest Approach: 

```{r}
dat_US <- dat[dat$Country.Region == "US",]
trCtl <- trainControl(method="cv", number=5, savePredictions=TRUE)
tg <- data.frame(mtry=1)
rf_fit <- train(dat_US["date"], dat_US$new_cases, method="rf", trControl=trCtl, tuneGrid=tg) 

pred <- rf_fit$pred[order(rf_fit$pred$rowIndex),]

with(dat_US, plot(dat_US["date"], dat_US$new_cases))
points(dat_US$date, pred$pred, col="red")
```

I don't know how well we can do prediction with this? It's not like we get parameter estimates? 

But I wrote a function here to get the graphs & predictions for a given country:
```{r}
randomForest_country <- function(country, data){
  trCtl <- trainControl(method="cv", number=5, savePredictions=TRUE)
  tg <- data.frame(mtry=1)
  
  
  dat <- data[data$Country.Region == country, ]
  rf_fit <- train(dat["date"], dat$new_cases, method="rf", trControl=trCtl, tuneGrid=tg) 
  pred <- rf_fit$pred[order(rf_fit$pred$rowIndex),]
  
  with(dat, plot(dat["date"], dat$new_cases))
  points(dat$date, pred$pred, col="red")
  
  return(pred)
}
```

```{r}
US_rf <- randomForest_country("US", dat)
# China_rf <- randomForest_country("China", dat)
SKorea_rf <- randomForest_country("Korea, South", dat)
Italy_rf <- randomForest_country("Italy", dat)
```
